# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
First let's create some new simpleDate class to parse the csv

```{r, echo=TRUE}
library(ggplot2)
setClass("simpleDate")
setAs("character","simpleDate", function(from) as.Date(from, format="%Y-%m-%d") )

```
First, we are unzipping the file to the *zip* variable
Once it's unzip, we can load the data to a new variable 
```{r, echo=TRUE}

zip<-unz("activity.zip",filename="activity.csv")
stepAct <- read.csv(zip,  sep=",", na.strings=c("", " ", "?"), header=TRUE, colClasses=c("numeric", "simpleDate", "numeric"))
```
Let's check the summary of our dataset
```{r, echo=TRUE}
summary(stepAct)
```

We'll have a first look to the number of steps per day with an histogram

```{r, echo=TRUE}
stepsPerDay <- aggregate(steps~date, stepAct, FUN=sum)
ggplot(data=stepsPerDay, aes(x=steps)) + geom_histogram(colour="black", fill="grey", binwidth=1000)
```
## What is mean total number of steps taken per day?

```{r, echo=TRUE}
paste("The mean is : ", mean(stepsPerDay$steps, na.rm=TRUE ))
paste("The median is : ", median(stepsPerDay$steps, na.rm=TRUE ))
```

## What is the average daily activity pattern?

```{r, echo=TRUE}
meanStepsPerInterval <- aggregate(steps~interval, stepAct, FUN=mean)
summary(meanStepsPerInterval)
ggplot(meanStepsPerInterval, aes(x=interval, y=steps))+
  geom_line(color="darkred", size=1) 
```
```{r, echo=TRUE}
paste("The interval with the max number of steps is : ",meanStepsPerInterval[which.max(meanStepsPerInterval$steps),]$interval )
```

## Imputing missing values
For imputing the missing values, we will take the mean of the corresponding 5-minutes
interval.  
The merge function is used to merge the two datasets on interval.
```{r, echo=TRUE}

stepActImputed <- stepAct

rM  = merge(stepActImputed, meanStepsPerInterval, by="interval", suffixes=c("orig", "mean"))
rMna.idx = which(is.na(stepActImputed$steps))
nbNa <- length(rMna.idx) 
```
paste("The number of NA values is  : ",nbNa)


```{r, echo=TRUE}
stepActImputed[rMna.idx,"steps"] = rM[rMna.idx,"stepsmean"]
summary(stepActImputed)

meanStepsPerIntervalImputed <- aggregate(steps~date, stepActImputed, FUN=sum)
ggplot(data=meanStepsPerIntervalImputed, aes(x=steps)) + geom_histogram(colour="black", fill="grey", binwidth=1000)
paste("The mean  is : ", mean(meanStepsPerIntervalImputed$steps))
paste("The median is : ", median(meanStepsPerIntervalImputed$steps))
```
These new values **do** increase the mean and median, so not only they are different but they also increased their values.
## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
stepActImputed$weekDays <- as.factor(ifelse(weekdays(stepActImputed$date) %in% c("Saturday","Sunday"), "Weekend", "Weekday")) 
meanWeekDaysActImputed <- aggregate(steps~interval+ weekDays, data=stepActImputed, mean)
ggplot(data=meanWeekDaysActImputed, 
       aes(x=interval, y=steps)) + 
  geom_line(color="darkred", size=1) + 
  facet_wrap(~weekDays, nrow=2, ncol=1)

```

As we can see with the plots, there are **some differences between the two graphs**.  
During the week days, in the **earlier** moment of the days, we can see that the number of steps are higher, most likely due to people waking up to go to work.  
During the weekdays/daytime, we can assume that people are **less active** as they are working until they get off from work.  
During the week end, the average is more **spread** accross the day.